---
title: "07072026_inclass"
author: "Arabella"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
coronavirus <- read.csv('https://raw.githubusercontent.com/RamiKrispin/coronavirus/master/csv/coronavirus.csv')
```
From lesson 9: Let’s start with summarizing the total number of cases by type as of the most recent day in the dataset: 2023-01-04. Take a minute to try this for yourself, then you can look at our approach.
```{r}
names(coronavirus)
cvtotals <- coronavirus %>% group_by(type) %>% 
  summarize(cases=sum(cases))
kable(cvtotals)
```
Now, let’s plot the history of the daily counts of new confirmed cases worldwide
```{r}
coronavirus %>%  
  filter(type == "confirmed") %>% 
  group_by(date) %>% 
  summarize(cvtotals = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = cvtotals)) +
  geom_line()
```
Let's riff on that with the same data and make a new graph 
```{r}
basegraph <- coronavirus %>%  
  filter(type == "confirmed") %>% 
  group_by(date) %>% 
  summarize(cvtotals = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = cvtotals)) 
```
```{r}
basegraph + 
  geom_point(
    aes(size = cvtotals,
        color = cvtotals),
    alpha = 0.5
  ) + 
  theme_minimal()

```
Updating axis labels and titles: use labs to update axis labels
```{r}
basegraph +
  geom_line() +
  theme_classic() +
  labs(
    x="Date", 
    y="Total confirmed cases",
    subtitle = "Global sums")
  
```
Take a minute to think about how you would generate a plot with a separate line showing the daily reports of new confirmed cases in each country.
```{r}
coronavirus %>% 
  filter(type == "confirmed") %>% 
  group_by(country, date) %>% 
  summarise(total_cases = sum(cases)) %>% 
  ggplot(mapping=(aes(x = date, y = total_cases, color = country))) + 
  geom_line()
```
This does not display very well, so try laying down parameters. In this example, we're subsetting 10 countries with the highest total counts of confirmed cases. 

```{r}
top10_countries <- coronavirus %>% 
  filter(type == "confirmed") %>% 
  group_by(country) %>% 
  summarize(cvtotals = sum(cases)) %>% 
  arrange(-cvtotals) %>% 
  head(10) %>% 
  .$country


coronavirus %>% 
  filter(type =="confirmed", country %in% top10_countries) %>% 
  group_by(country, date) %>% 
  summarize(total_cases = sum(cases)) %>% 
  ggplot(mapping = aes(x = date, y = total_cases, color = country)) + geom_line()+
  facet_wrap(~country)
```
```{r}
coronavirus %>% 
  filter(type =="confirmed", country %in% top10_countries) %>% 
  group_by(country) %>% 
  arrange(date) %>%
  mutate(cum_count = cumsum(cases)) %>%
  ggplot(mapping=aes(x=date, y=cum_count, color=country)) + 
  geom_line()
```
Time for bar charts!

```{r}
No_negatives <- coronavirus %>% filter(cases >=0)
No_negatives %>%  
  group_by(date) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x = date, y = cases), color = "black")
  
```
```{r}
No_negatives %>%  
  group_by(date, type) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x = date, y = cases, fill=type), size=.01)
```
```{r}
No_negatives %>%  
  group_by(date, type) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x = date, y = cases, fill=type), size=.01)+
  coord_flip()
```

```{r}
coronavirus %>% 
  group_by(date, type) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x=date, y = cases, fill = type), color="black", size=0.1, position = "fill")
```
```{r}
coronavirus %>% 
  filter(cases > 0) %>% 
  group_by(date, type) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x = date, y = cases, fill = type), size=0.1, position = "fill")
```
```{r}
coronavirus %>% 
  filter(cases > 0) %>% 
  group_by(date, type) %>%
  summarize(cases = sum(cases)) %>%
  ggplot() +
  geom_col(aes(x = date, y = cases, fill = type), position = "dodge")
```
```{r}
install.packages("rnaturalearth")
install.packages("rnaturalearthdata")
install.packages("sf")
```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
filter(coronavirus, date == max(coronavirus$date), type == "confirmed", cases > 5000) %>%
  ggplot() +
  geom_sf(data = world) +
  geom_point(aes(x=long, y=lat, size=cases), color="red", fill="red", alpha=0.5, shape=21) 
```


